EoS-insensitive relations after GW170817

### 2019-02-11 paper/bin-love-after-gw170817.tex
implemented comments from teleconference, and also did a thourough read through to fix flow, grammer, english, spelling,, and general issues

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Begin Abstract %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% Intro
The thermodynamic relationship between pressure and density (the equation of state, or EoS) of cold supranuclear matter, accessible in neutron stars, is critical to the study of such stars, and is one of the largest uncertainties in nuclear physics to date. 
% How to probe
The extraction of tidal deformabilities from the gravitational waveforms of binary neutron star merger events, such as GW170817, is a promising method of probing the nuclear structure.
% What was done before - Binary Love
Previous studies have shown that approximately EoS-insensitive ``binary Love relations" exist between symmetric and antisymmetric combinations of individual tidal deformabilities, which are difficult to be independently measured with second-generation gravitational wave interferometers.
% What was done before - I-Love-Q
Similarly, another set of EoS-insensitive relations exist between individual neutron star parameters: moment of inertia (I), tidal deformability (Love number), quadrupole moment (Q), and compactness (C) - known as the ``I-Love-Q" and ``C-Love" relations.
% What universal relations do
Such EoS-insensitive relations allow the elimination of some tidal parameters from the list of model parameters, thus breaking degeneracies and ultimately reducing parameter extraction uncertainties.
% What we will do
However, even the most precise of EoS insensitive relations carry EoS variation which inject systematic errors into the parameter extraction process.
In this document, we explain how one can reduce equation of state variation for I-Love-Q, C-Love, and binary Love relations, which helps to reduce systematic errors on the tidal measurement of binary neutron star mergers.
However, we show that the systematic effects of these relations is very small until next-generation detectors. 
We achieve this  by restricting to only equations of state drawn from the 90\% posterior constraint on pressure as a function of density, as derived by the LIGO Collaboration.
% What we found
We find an improvement in binary Love universality by a factor of $\sim 59$\% for stars with a mass ratio of 0.75, in C-Love universality by a factor of $\sim 67\%$, and in I-Love-Q universality by factors of $\sim 50$\%.
% Conclude
We end by comparing systematic errors on tidal measurement due to the equation of state variation with statistical errors and comment on whether one can safely use these EoS-insensitive relations with future gravitational wave observations.
We conclude that the systematic errors injected from the use of improved EoS-insensitive relations become comparable to the statistical errors for future design upgrade Voyager.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Begin Introduction %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}\label{sec:intro}
%Intro

{\ny{Just some general comments so that we don't forget:
\begin{itemize}
\item The EoS-insensitive relations are called ``binary Love'', ``I-Love-Q'', ``C-Love'' and ``R-Love.'' There is no need to call them ``binary Love EoS-insensitive relations.'' That's too verbose and not what they are called in the literature. I am ok with describing these relations as EoS-insensitive instead of approximately universal, but let's not change the names. 
\item There is some serious tidying up that needs to be done to all of the text. Specifically, there are missing prepositions, like ``the'', in lots of places that makes the reading a bit choppy and we can improve. Zack, when you have a chance, please do a detail read through and fix the flow as much as you can. 
\item Eventually, we need to discuss the abstract and the intro in more detail. 
\item Let's not forget to check the acknowledgments for funding sources. 
\item Let's do a spell check. 
\item Let's remove all commented out text.  
\end{itemize}
}}
The dependence of pressure on density for cold supranuclear matter found inside neutron stars (NSs) remains to be one of the largest mysteries in both nuclear physics and astrophysics.
This internal structure of such stars, known as the equation of state (EoS), is exceedingly important, as it is the determining factor for many NS observables - such as the mass, the radius, and many more.
Unfortunately, terrestrial experiments can only study the EoS to around the nuclear saturation density ($\rho_{\text{sat}} \approx 2.5 \times 10^{14} \text{ g/cm}^3$)~\cite{Li:HeavyIon,Tsang:SymmetryEnergy,Centelles:NeutronSkin,Li:CrossSections,Chen:SymEnergy} (although some temperature-dependent heavy-ion collision experiments may probe higher densities~\cite{Danielewicz:2002pu}), making NSs ideal laboratories for constraining ultra dense nuclear matter.

%Constraining the EoS
Independent measurements of NS observables can be used to constrain the nuclear EoS.
For example, electromagnetic observations of the mass-radius relationship of NSs have been used to place limits on the NS EoS~\cite{guver,ozel-baym-guver,steiner-lattimer-brown,Lattimer2014,Ozel:2016oaf}.
However, these implications potentially suffer from large systematic errors due to various x-ray burst astrophysical mismodeling uncertainties.
Alternatively, the emission of gravitational waves (GWs) from binary NS merger events have proven to be a unique method of probing nuclear physics.
During the early inspiral portion of the merger, the orbital separation is large enough that the effect of companion tidal fields is negligible. 
As the separation decreases through GW emission, tidal forces are magnified and the NSs deform from sphericity, altering the orbital trajectories; a relic which is directly imprinted on the GW signal.
This deformation is characterized by the \textit{tidal deformability} $\Lambda$ as the linear response of the NSs quadrupole moment $Q_{ij}$ to the neighboring tidal field $\varepsilon_{ij}$~\cite{hinderer-love,Flanagan2008}.

%reparameterization of the template
In the case of binary NSs, each star becomes deformed via the process highlighted above, resulting in two highly correlated tidal parameters $\Lambda_1$ and $\Lambda_2$ entering in the gravitational waveform\cite{Flanagan2008,Vines:2011ud}.
Due to high correlations between between the two parameters, independent extraction becomes very difficult at current interferometer sensitivities~\cite{Wade:tidalCorrections}.
Typically, these high correlations can be reduced by strategically reparameterizing the waveform template with new tidal parameters constructed from linear combinations of $\Lambda_1$ and $\Lambda_2$.
For example, utilizing the parameters $\tilde{\Lambda}$ and $\delta \tilde{\Lambda}$~\cite{Favata:2013rwa,Wade:tidalCorrections} which enter the GW waveform at different post-Newtonian (PN) orders (powers of the ratio between the orbital speed and the speed of light $v/c^{2n}$) partially mitigates these correlations.
Similarly, the waveform may be parameterized by mass-independent tidal parameters $\lambda_0$ and $\lambda_1$, derived through a Taylor expansion about $m_0=1.4 \text{ M}_{\odot}$ of tidal deformability $\Lambda \approx \lambda_0+\lambda+1(1-\frac{m}{m_0})$~\cite{delPozzo:TaylorTidal,Yagi:binLove}. 
Use of this set of tidal parameters allows one to effectively combine multiple events with varying mass and $\tilde\Lambda$ components, not possible in the former parameterization of $\tilde{\Lambda}$ and $\delta \tilde{\Lambda}$

%Universal relations
Unfortunately, current detectors are only sensitive enough to accurately measure the dominant tidal parameter, $\lambda_0$ (also known as $\Lambda_{1.4}$, the tidal deformability at $1.4 \text{ M}_{\odot}$).
Previous work by Yagi and Yunes~\cite{Yagi:binLove} resolved this issue by finding approximately EoS-insensitive ``binary Love relations" between symmetric and anti-symmetric combinations of tidal deformabilities $\Lambda_{s,a}=\frac{1}{2}(\Lambda_1 \pm \Lambda_2)$.
This allows one to further break degeneracies between tidal parameters, resulting in both (i) a decrease in uncertainty upon parameter extraction of $\Lambda_s$ on the removal of parameter $\Lambda_a$ from the model, and (ii) the ability to measure the secondary tidal parameter upon measurement of the first~\cite{Katerina:residuals}.  
Use of these relations leads to up to an order of magnitude improvement in parameter estimation of $\Lambda_s$ through a simple Fisher analysis~\cite{Yagi:binLove}, also confirmed with the detailed analysis of~\cite{Katerina:residuals}.
Similar ``I-Love-Q" and ``C-Love" relations~\cite{Yagi:ILQ} have been found between individual NS observables: quadrupole moment , moment of inertia, tidal deformability, and compactness; similarly allowing the removal of tidal parameters from the model list, and allowing an automatic determination of one observable through the measurement of another.
However, these relations have been tuned to a collection of EoS models, and each have intrinsic uncertainties associated with their use.

%How we improved this
In this analysis, we aim to improve this important work by imposing the restrictions found on the EoS~\cite{LIGO:posterior} derived from the recent binary NS merger observation, GW170817~\cite{TheLIGOScientific:2017qsa}.
By restricting to only EoSs which lie within the the 90\% credible region on pressure as a function of density found in~\cite{LIGO:posterior}, we repeat the analyses done in~\cite{Yagi:binLove,Yagi:ILQ} and show an increase in universality for I-Love-Q, C-Love, and binary Love relations.
We also investigate the relationship between the NS radius and its tidal deformability, or the R-Love relations, determining if the uncertainties can be reduced.
In addition, we consider hybrid star EoSs seen in~\cite{Paschalidis2018}, which experience strong first-order transitions from hadronic matter to quark matter.
These provide a departure from the nuclear matter relations considered here - providing valuable insight into them.
Further, we estimate the viability of using improved EoS-insensitive relations on future GW detections by performing simple Fisher analyses to approximate when the statistical errors from parameter extraction become comparable to the systematic errors due to EoS variation in the relations.
To estimate the effect of waveform mismodeling, we additionally consider two different tidal corrections to the injected PhenomD~\cite{PhenomDI,PhenomDII} waveform: 6PN~\cite{Wade:tidalCorrections} and NRTidal~\cite{Samajdar:NRTidal}

\subsection{Executive Summary}
%What we did
In this paper we attempt to improve I-Love-Q, C-Love, and binary Love relations by imposing constraints on the EoS found by GW170817~\cite{LIGO:posterior,TheLIGOScientific:2017qsa}.
First, we generate two large samples of spectral EoSs~\cite{Lindblom:2018rfr}, where we (i) impose the restriction that they must be drawn from GW170817s 90\% credible posterior in pressure as a function of density, and (ii) don't impose any restrictions for comparison purposes.
Next, we follow the important works of~\cite{Yagi:binLove} and~\cite{Yagi:ILQ} to show how the ``constrained" set of EoSs show improved universality (binary Love, I-Love-Q, and C-Love) from both the previous works, and the ``unconstrained" sample, shown by Fig.~\ref{fig:ILQ}.
In addition, we determine the degree to which hybrid star EoSs obey binary Love relations.
Finally, through a simple Fisher analysis, we consider the value in improving EoS-insensitive relations by approximating when the statistical errors from parameter extraction become comparable to the systematic errors due to EoS variation in the relations.

%Final Results
Upon use of the constrained set of EoSs, we find an improvement in binary Love universality by a factor of $\sim 59$\% for stars with mass ratios of $0.75$, in C-Love universality by a factor of $\sim 67\%$, and in I-Love-Q universality by factors of $\sim50$\%, tabulated in Tab.~\ref{tab:maxVar}.
We also present the relationship between the NS radius and tidal deformability, computed directly from the C-Love relations, showing maximum EoS variability of $358.9 \text{ km}$ from the fit discussed in Sec.~\ref{sec:clove}.
Additionally, we find that hybrid quark-hadron stars do obey the I-Love-Q and C-Love relations, with sightly increased EoS variability.
We also find that binaries consisting of one massive hybrid star, and one small-mass hadronic star do not agree with derived binary Love relations, due to the large separation in $\tilde{\Lambda}$ between the constituent stars.
Due to the current limitations in detector sensitivity, the systematic errors arising from EoS variation in the EoS-insensitive relations are far outweighed by the statistical errors accrued by extraction of high-order PN tidal parameters from GW waveforms.
Fig.~\ref{fig:stackedFisher} compiles the results from Fisher analyses approximating the uncertainties accrued from the parameter extraction of $\lambda_0$ (the tidal deformability at $1.4 \text{ M}_{\odot}$) for GW170817 as if it had been observed by future detectors $A \equiv ($ O2~\cite{aLIGO}, aLIGO~\cite{aLIGO}, A\texttt{+}~\cite{Ap_Voyager_CE}, Voyager~\cite{Ap_Voyager_CE}, ET~\cite{ET}, CE~\cite{Ap_Voyager_CE}$)$.
Here, $\sigma^A_{\text{GW170817}}$ and $\rho^A_{\text{GW170817}}$ correspond to the statistical error and signal-to-noise-ratio found from observing GW170817 on detector $A$, and $\sigma^A_N$ represents the combined uncertainty after the detection of $N_A$ events (corresponding to the binary NS merger rate associated interferometer $A$).
These values are further tabulated in Table~\ref{tab:variances}.
It can be seen from this figure that the statistical errors from parameter extraction become comparable the EoS-insensitive relation systematics (indicated by the magenta horizontal line) for the Voyager telescope - indicating when improved binary Love relations will become necessary.
Also shown in the figure is the same measurement uncertainties upon the injection of the PhenomD + NRTidal waveform, rather than the PhenomD + 6PN tidal correction to demonstrate the effect of waveform mismodeling.
\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{stackedFisher.eps}
\end{center}
\caption{
Estimated statistical uncertainties on the extraction of the tidal deformability at $1.4 \text{ M}_{\odot}$, $\lambda_0$, observed from interferometers $A=$(O2, aLIGO, A\texttt{+}, Voyager, CE, ET-D$)$ as a function of the signal-to-noise-ratio.
In this figure, the filled blue circles denote the single-event statistical uncertainties $\sigma^A_{\text{GW170817}}$ of GW170817 as if it were observed on detector $A$.
Also shown by the turquoise shaded region is combined uncertainty $\sigma^A_N$  from the observation of multiple events from a simulated population of size $N_A$ corresponding to the upper and lower limits of the expected binary NS merger detection rates for O2, aLIGO, A\texttt{+}, Voyager, CE, and ET-D over 1 observation year.
It can be seen that the statistical errors in the parameter extraction become comparable to the systematics from EoS variation in the improved binary Love relations (indicated by the horizontal indigo dashed line) for Voyager, CE, and ET.
These results are further tabulated in Table~\ref{tab:variances}.
Additionally shown in the figure are the estimated statistical uncertainties on the extraction of $\lambda_0$ upon the injection of the PhenomD + NRTidal correction waveform (IMRD + NRTidal), rather than the PhenomD + 6PN tidal correction (IMRD + 6PN), in order to demonstrate the effect of waveform mismodeling uncertainties.
The single-detection uncertainties are shown by unfilled maroon circles, while the combined detection uncertainties are outlined by dashed maroon.
}
\label{fig:stackedFisher}
\end{figure}

%outline
The organization of this paper is outlined below.
We begin with a complementary background and theory material in Sec.~\ref{sec:theory}.
We continue in Sec.~\ref{sec:universal} by finding new and improved binary Love, I-Love-Q, and C-Love relations, and considering how well hybrid star EoSs agree with the EoS-insensitive relations.
We next examine these improved relations and question whether or not they are useful for future interferometers in Sec.~\ref{sec:observations}.
We conclude in Sec.~\ref{sec:conclusion} by discussing our results and mentioning avenues of future work.
Throughout this paper, we have adopted geometric units of $G=c=1$, unless otherwise stated.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Begin theory %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Background and theory}\label{sec:theory}

\subsection{Spectral representations of neutron star equations of state}\label{sec:eos}
The structure of a NS and its tidal interactions in a binary system rely heavily on the underlying state function (or equation of state) describing the relationship between pressure ($p$) and energy density ($\epsilon$) of nuclear matter.
Given that all currently proposed EoSs utilize certain approximations~\cite{Oertel:Review,Baym:Review}, one method to study a wide range of physically realizable EoSs is to parameterize them such that any realistic EoS can be represented with a small number of parameters.
Spectral representations~\cite{Lindblom:2010bb,Lindblom:2012zi,Lindblom:2013kra,Lindblom:2018rfr,Abbott:2018exr} parameterize EoSs by performing spectral expansions on the adiabatic index $\Gamma(p)$\footnote{Another way of parameterizing EoSs is the piecewise polytropic formulation~\cite{Read2009,Lackey:2014fwa,Carney:2018sdv}.}:
\begin{equation}
\Gamma(x) = \exp{\sum_k\gamma_k x^k},
\end{equation}
where $x \equiv \log{p/p_0}$ for a minimum pressure $p_0$.
The equation of state is then determined by an integration of the differential equation:
\begin{equation}
\frac{d \epsilon(p)}{dp}=\frac{\epsilon(p)+p}{p \Gamma(p)}.
\end{equation}
Using this formalism, any valid EoS can be approximated through the choice of 4 or more spectral coefficients $\gamma_k$, tabulated for several common EoSs in Table 1 of~\cite{Lindblom:2018rfr}.

>>>>

%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%
%%%% ABSTRACT %%%%
%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%

\begin{abstract}
% Intro
The thermodynamic relationship between pressure and density (the equation of state, or EoS) of cold supranuclear matter, accessible in neutron stars, is critical to the study of such stars, and is one of the largest uncertainties in nuclear physics to date. 
% How to probe
The extraction of tidal deformabilities from the gravitational waveforms of binary neutron star merger events, such as GW170817, is a promising method of probing the nuclear structure.
% What was done before - Binary Love
Previous studies have shown that approximately EoS-insensitive ``binary Love relations" exist between symmetric and antisymmetric combinations of individual tidal deformabilities, which are difficult to be independently measured with second-generation gravitational wave interferometers.
% What was done before - I-Love-Q
Similarly, another set of EoS-insensitive relations exist between individual neutron star parameters: moment of inertia (I), tidal deformability (Love number), quadrupole moment (Q), and compactness (C) - known as the ``I-Love-Q" and ``C-Love" relations.
% What universal relations do
Such EoS-insensitive relations allow the elimination of some tidal parameters from the list of model parameters, thus breaking degeneracies and ultimately reducing the parameter extraction uncertainties.
% What we will do
However, even the most precise of EoS-insensitive relations carry EoS variation uncertainties which inject systematic errors into the parameter extraction process.
In this document, we explain how one can reduce the equation of state variation for the I-Love-Q, C-Love, and binary Love relations, which helps to reduce the systematic errors on the tidal measurement of binary neutron star mergers.
However, we show that the systematic effects of these relations are negligible until next-generation detector sensitivities. 
We achieve this  by restricting to only equations of state drawn from the 90\% posterior constraint on pressure as a function of density, as derived by the LIGO Collaboration.
% What we found
We find an improvement in binary Love universality by a factor of $\sim 59$\% for stars with a mass ratio of 0.75, in C-Love universality by a factor of $\sim 67\%$, and in I-Love-Q universality by factors of $\sim 50$\%.
% Conclude
We end by comparing the systematic errors on tidal measurements due to the equation of state variation with statistical errors and comment on whether one can safely use these EoS-insensitive relations with future gravitational wave observations.
We conclude that the systematic errors injected from the use of improved EoS-insensitive relations become comparable to the statistical errors for future detector upgrade Voyager.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%
%%%% INTRODUCTION %%%%
%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}\label{sec:intro}

{\ny{Just some general comments so that we don't forget:
\begin{itemize}
\item The EoS-insensitive relations are called ``binary Love'', ``I-Love-Q'', ``C-Love'' and ``R-Love.'' There is no need to call them ``binary Love EoS-insensitive relations.'' That's too verbose and not what they are called in the literature. I am ok with describing these relations as EoS-insensitive instead of approximately universal, but let's not change the names. 
\item There is some serious tidying up that needs to be done to all of the text. Specifically, there are missing prepositions, like ``the'', in lots of places that makes the reading a bit choppy and we can improve. Zack, when you have a chance, please do a detail read through and fix the flow as much as you can. 
\item Eventually, we need to discuss the abstract and the intro in more detail. 
\item Let's not forget to check the acknowledgments for funding sources. 
\item Let's do a spell check. 
\item Let's remove all commented out text.  
\end{itemize}
}}
%Intro
The dependence of pressure on density for cold supranuclear matter found inside neutron stars (NSs) remains to be one of the largest mysteries in both nuclear physics and astrophysics.
This internal structure of such stars, known as the equation of state (EoS), is exceedingly important, as it is the determining factor for many NS observables - such as the mass and the radius.
Unfortunately, terrestrial experiments can only study the EoS to around the nuclear saturation density ($\rho_{\text{sat}} \approx 2.5 \times 10^{14} \text{ g/cm}^3$)~\cite{Li:HeavyIon,Tsang:SymmetryEnergy,Centelles:NeutronSkin,Li:CrossSections,Chen:SymEnergy} (although some temperature-dependent heavy-ion collision experiments may probe higher densities~\cite{Danielewicz:2002pu}), making NSs ideal laboratories for constraining ultra dense nuclear matter.

%Constraining the EoS
Independent measurements of NS observables can be used to constrain the nuclear EoS.
For example, electromagnetic observations of the NS mass-radius relationship have been used to place limits on the EoS~\cite{guver,ozel-baym-guver,steiner-lattimer-brown,Lattimer2014,Ozel:2016oaf}.
However, these implications potentially suffer from large systematic errors due to various x-ray burst astrophysical mismodeling uncertainties.
Alternatively, the emission of gravitational waves (GWs) from binary NS merger events have proven to be a unique method of probing nuclear physics.
During the early inspiral portion of the merger, the orbital separation is large enough that the effect of companion tidal fields is negligible. 
As the separation decreases through GW emission, tidal forces are magnified and the NSs deform from sphericity, altering the orbital trajectories; a relic which is directly imprinted on the GW signal.
This deformation is characterized by the \textit{tidal deformability} $\Lambda$ as the linear response of the NSs quadrupole moment $Q_{ij}$ to the neighboring tidal field $\varepsilon_{ij}$~\cite{hinderer-love,Flanagan2008}.

%reparameterization of the template
In the case of binary NSs, each star becomes deformed via the process highlighted above, resulting in two highly correlated tidal parameters $\Lambda_1$ and $\Lambda_2$ entering the gravitational waveform~\cite{Flanagan2008,Vines:2011ud}.
Due to high correlations between between the two parameters, independent extraction becomes very difficult at current interferometer sensitivities~\cite{Wade:tidalCorrections}.
Typically, these high correlations can be reduced by strategically reparameterizing the waveform template with new tidal parameters constructed from linear combinations of $\Lambda_1$ and $\Lambda_2$.
For example, utilizing the parameters $\tilde{\Lambda}$ and $\delta \tilde{\Lambda}$~\cite{Favata:2013rwa,Wade:tidalCorrections} which enter the GW waveform at different post-Newtonian (pN) orders\footnote{n-pN order corrections enter the gravitational waveform at powers of the ratio between the orbital speed and the speed of light $v/c^{2n}$.} partially mitigates these correlations.
Similarly, the waveform may be parameterized by mass-independent tidal parameters $\lambda_0$ and $\lambda_1$, derived through a Taylor expansion about $m_0=1.4 \text{ M}_{\odot}$ of tidal deformability $\Lambda \approx \lambda_0+\lambda_1(1-\frac{m}{m_0})$~\cite{delPozzo:TaylorTidal,Yagi:binLove}. 
Use of this set of tidal parameters allows one to effectively combine multiple events with varying mass and $\tilde\Lambda$ components, not possible in the former parameterization of $\tilde{\Lambda}$ and $\delta \tilde{\Lambda}$

%Universal relations
Unfortunately, current detectors are not sensitive enough to accurately measure the dominant tidal parameter, $\lambda_0$ (also known as $\Lambda_{1.4}$, the tidal deformability at $1.4 \text{ M}_{\odot}$).
Previous work by Yagi and Yunes~\cite{Yagi:binLove} resolved this issue by finding approximately EoS-insensitive ``binary Love relations" between symmetric and anti-symmetric combinations of tidal deformabilities $\Lambda_{s,a}=\frac{1}{2}(\Lambda_1 \pm \Lambda_2)$.
This allows one to further break degeneracies between tidal parameters, resulting in both (i) a decrease in uncertainty upon parameter extraction of $\Lambda_s$ on the removal of parameter $\Lambda_a$ from the model; and (ii) the ability to measure the secondary tidal parameter upon measurement of the first~\cite{Katerina:residuals}.  
Use of these relations leads to up to an order of magnitude improvement in parameter estimation of $\Lambda_s$ through a simple Fisher analysis~\cite{Yagi:binLove}, also confirmed with the detailed analysis seen in Ref.~\cite{Katerina:residuals}.
Similar ``I-Love-Q" and ``C-Love" relations~\cite{Yagi:ILQ} have been found between individual NS observables: quadrupole moment, moment of inertia, tidal deformability, and compactness; similarly allowing the removal of tidal parameters from the model list, and allowing an automatic determination of one observable through the measurement of another.
However, these relations have been tuned to a specific collection of EoS models, and each have intrinsic uncertainties associated with their use.

%How we improved this
In this analysis, we aim to improve this important work by imposing the restrictions found on the EoS~\cite{LIGO:posterior} derived from the recent binary NS merger observation, GW170817~\cite{TheLIGOScientific:2017qsa}.
By restricting to only EoSs which lie within the the 90\% credible region on pressure as a function of density found in Ref.~\cite{LIGO:posterior}, we repeat the analyses done in Refs.~\cite{Yagi:binLove,Yagi:ILQ} and show an increase in universality for the I-Love-Q, C-Love, and binary Love relations.
We also investigate the relationship between the NS radius and its tidal deformability, or the R-Love relations, determining whether the NS radius uncertainties can be reduced.
In addition, we consider hybrid star EoSs seen in Ref.~\cite{Paschalidis2018}, which experience strong first-order transitions from hadronic matter to quark matter in the core.
These provide a departure from the nuclear matter relations considered here - providing valuable insight into such EoSs.
Lastly, we estimate the viability of using the improved EoS-insensitive relations on future GW detections by performing simple Fisher analyses to approximate when the statistical errors from parameter extraction become comparable to the systematic errors due to EoS variation in the relations.
To further estimate the effect of waveform mismodeling, we additionally consider two different tidal corrections to the injected PhenomD~\cite{PhenomDI,PhenomDII} waveform: 6PN~\cite{Wade:tidalCorrections} and NRTidal~\cite{Samajdar:NRTidal}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% EXECUTIVE SUMMARY %%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Executive Summary}
%What we did
In this paper we attempt to improve the I-Love-Q, C-Love, and binary Love relations by imposing constraints on the EoS found by GW170817~\cite{LIGO:posterior,TheLIGOScientific:2017qsa}.
First, we generate two large samples of spectral EoSs~\cite{Lindblom:2018rfr}, where we (i) impose the restriction that they must be drawn from GW170817s 90\% credible posterior in pressure as a function of density; and (ii) don't impose any restrictions for comparison purposes.
Next, we follow the important works of Refs.~\cite{Yagi:binLove} and~\cite{Yagi:ILQ} to show how the ``constrained" set of EoSs show improved universality (binary Love, I-Love-Q, and C-Love) from both the previous works, and the ``unconstrained" sample, shown by Fig.~\ref{fig:ILQ}.
In addition, we determine the degree to which hybrid star EoSs obey the various EoS-insensitive relations.
Finally, through a simple Fisher analysis, we consider the value in improving EoS-insensitive relations by approximating when the statistical errors from parameter extraction become comparable to the systematic errors due to EoS variation in the relations.

%Final Results
Upon use of the constrained set of EoSs, we find an improvement in binary Love universality by a factor of $\sim 59$\% for stars with mass ratios of $0.75$, in C-Love universality by a factor of $\sim 67\%$, and in I-Love-Q universality by factors of $\sim50$\%, tabulated in Tab.~\ref{tab:maxVar}.
We also present the relationship between the NS radius and tidal deformability, computed directly from the C-Love relations, showing maximum EoS variability of $358.9 \text{ km}$ from the fit discussed in Sec.~\ref{sec:clove}.
Additionally, we find that the hybrid quark-hadron star EoSs do obey the I-Love-Q and C-Love relations, with sightly increased EoS variability.
We also find that binaries consisting of one massive hybrid star, and one small-mass hadronic star do not agree with the derived binary Love relations, due to the large separation in $\tilde{\Lambda}$ between the constituent stars.

Due to the current limitations in detector sensitivity, the systematic errors arising from EoS variation in the EoS-insensitive relations are far outweighed by the statistical errors accrued by the extraction of high pN order tidal parameters from GW waveforms.
Fig.~\ref{fig:stackedFisher} compiles the results from Fisher analyses approximating the uncertainties accrued from the parameter extraction of $\lambda_0$ (the tidal deformability at $1.4 \text{ M}_{\odot}$) for GW170817 as if it had been observed by various detectors $A \equiv ($ O2~\cite{aLIGO}, aLIGO~\cite{aLIGO}, A\texttt{+}~\cite{Ap_Voyager_CE}, Voyager~\cite{Ap_Voyager_CE}, ET~\cite{ET}, CE~\cite{Ap_Voyager_CE}$)$.
Here, $\sigma^A_{\text{GW170817}}$ and $\rho^A_{\text{GW170817}}$ correspond to the statistical error and signal-to-noise-ratio found from observing GW170817 on detector $A$, and $\sigma^A_N$ represents the combined uncertainty after the detection of $N_A$ events (corresponding to the binary NS merger rate associated with interferometer $A$).
These values are further tabulated in Table~\ref{tab:variances}.
It can be seen from this figure that the statistical errors from parameter extraction become comparable the EoS-insensitive relation systematics (indicated by the magenta horizontal line) for the Voyager interferometer - indicating when the improved binary Love relations will become necessary.
Also shown in the figure is the same measurement uncertainties upon the injection of the PhenomD + NRTidal~\cite{PhenomDI,PhenomDII,Samajdar:NRTidal} waveform, rather than the PhenomD + 6PN~\cite{PhenomDI,PhenomDII,Wade:tidalCorrections} tidal correction to demonstrate the effect of waveform mismodeling.

\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{stackedFisher.eps}
\end{center}
\caption{
Estimated statistical uncertainties on the extraction of the tidal deformability at $1.4 \text{ M}_{\odot}$, $\lambda_0$, observed from interferometers $A=$(O2, aLIGO, A\texttt{+}, Voyager, CE, ET-D$)$ as a function of the signal-to-noise-ratio $\rho^{\text{A}}_{\text{GW170817}}$.
%evaluated from $300 \text{ Hz} \leq f \leq 1649 \text{ Hz}$\footnote{This was done for demonstrative purposes, as computing the SNR with $1 \text{ Hz} \leq f \leq 1649 \text{ Hz}$ resulted in smaller statistical uncertainties for ET than CE, albeit having a smaller SNR - a relic originating from ET being more sensitive to the high frequency tidal effects than CE.}.
In this figure, the filled blue circles denote the single-event statistical uncertainties $\sigma^A_{\text{GW170817}}$ of GW170817 as if it were observed on detector $A$\footnote{We note that ET more accurately measures single events than CE albeit having a smaller SNR - a relic originating from ET being more sensitive to the high frequency tidal effects than CE. See Sec.~\ref{sec:futureObservations} for a discussion on this.}.
Also shown by the cyan shaded region is the combined uncertainty $\sigma^A_N$ from the observation of multiple events from a simulated population of size $N_A$ corresponding to the upper and lower limits of the expected binary NS merger detection rates for detector $A$ over 1 observation year.
It can be seen that the statistical errors in the parameter extraction become comparable to the systematics from EoS variation in the improved binary Love relations (indicated by the horizontal indigo dashed line) for the Voyager, CE, and ET detectors.
These results are further tabulated in Table~\ref{tab:variances}.
Additionally shown in the figure are the estimated statistical uncertainties on the extraction of $\lambda_0$ upon the injection of the PhenomD + NRTidal correction waveform (IMRD + NRTidal), rather than the PhenomD + 6PN tidal correction (IMRD + 6PN), in order to demonstrate the effect of waveform mismodeling uncertainties.
The single-detection uncertainties are shown by unfilled maroon circles, while the combined detection uncertainties are outlined by dashed maroon.
}
\label{fig:stackedFisher}
\end{figure}

%outline
The organization of this paper is outlined below.
We begin with complementary background and theory material in Sec.~\ref{sec:theory}.
We continue in Sec.~\ref{sec:universal} by finding new and improved binary Love, I-Love-Q, and C-Love relations, and considering how well hybrid star EoSs agree with the EoS-insensitive relations.
We next examine these improved relations and question whether or not they are useful for future interferometers in Sec.~\ref{sec:observations}.
We conclude in Sec.~\ref{sec:conclusion} by discussing our results and mentioning avenues of future work.
Throughout this paper, we have adopted geometric units of $G=c=1$, unless otherwise stated.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% BACKGROUND AND THEORY %%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Background and theory}\label{sec:theory}

\subsection{Spectral representations of neutron star equations of state}\label{sec:eos}
The structure of a NS and its tidal interactions in a binary system rely heavily on the underlying state function (or equation of state - EoS) describing the relationship between the pressure ($p$) and energy density ($\epsilon$) of nuclear matter.
Given that all currently proposed EoSs utilize certain approximations~\cite{Oertel:Review,Baym:Review}, one method to study a wide range of physically realizable EoSs is to parameterize them such that any realistic EoS can be represented with a small number of parameters.
Spectral representations~\cite{Lindblom:2010bb,Lindblom:2012zi,Lindblom:2013kra,Lindblom:2018rfr,Abbott:2018exr} parameterize EoSs by performing spectral expansions on the adiabatic index $\Gamma(p)$\footnote{Another way of parameterizing EoSs is the piecewise polytropic formulation~\cite{Read2009,Lackey:2014fwa,Carney:2018sdv}.}:
\begin{equation}
\Gamma(x) = \exp{\sum_k\gamma_k x^k},
\end{equation}
where $x \equiv \log{p/p_0}$ for a minimum pressure $p_0$.
The equation of state is then determined by an integration of the differential equation:
\begin{equation}
\frac{d \epsilon(p)}{dp}=\frac{\epsilon(p)+p}{p \Gamma(p)}.
\end{equation}
Using this formalism, any valid EoS can be approximated through the choice of 4 spectral coefficients $\gamma_k$, tabulated for several common EoSs in Table 1 of Ref.~\cite{Lindblom:2018rfr}.
